Segmentation Using Attributes

### 2010-05-05 paper.tex
committed

We use the segmentation approach of \cite{arbelaez09}.  This
produces a hierarchical segmentation by first detecting contours
using the \emph{gPb} detector \cite{maire08}.  Once the contours are
detected, regions are discovered using a watershed algorithm after
which an Ultrametric Contour Map (UCM) is constructed, which defines
the region hierarchy. Some example of the hierarchy segmentation is
shown in Figure \ref{fig:tree}.

After the construction of hierarchy segmentation tree, we detect all
the attributes in each node, including the intermediate node and
leaf node. We use the output of SVM as the confidence (not binary),
so each node can be represented by the confidence distribution of
all attributes.

@@

Our objective is to detect object and segment it out. In our
hierarchy segmentation tree, the problem changes to how to select
those nodes belong to the desired object. We estimate the
probability that a node is belong to object based on the attribute
confidence distribution. For each object, we manually give the
desired attribute distribution by common knowledge. We use the
histogram intersection between the desired distribution and the
estimated confidence distribution in each node as the similarity
measure. The histogram intersection is calculated as follows:

@@

\begin{figure*}
\centering
\subfigure[]{
\label{fig:treea}
\includegraphics[width=0.45\textwidth]{figures/tree_additive.eps}
}
\subfigure[]{
\label{fig:treeb}
\includegraphics[width=0.45\textwidth]{figures/tree_subtractive.eps}
}
\caption{The two segmentation hierarchies used by the additive \subref{fig:treea}
and subtractive \subref{fig:treeb} algorithms.  The additive hierarchy breaks the
regions into smaller and smaller regions, while the subtractive hierarchy increases
the size of the regions as smaller and smaller chunks are removed from the image.}
\label{fig:tree}
\end{figure*}

@@

This pruning algorithm is based on three observations. (1) At least
one node in the segmentation tree contrains the whole object. (This
is always true since the root node contains the whole image). (2) If
both parent node and child node have high confidence of containing
the object, the segmentation of the child node is more accurate than
that of the parrent node. (3) Even if some node has high confidence
of containing the object, its segmentation can still be improved by
removing irrelevant children node. The details of the additive
algorithm is shown in algorithm \ref{alg:add_prune}.

@@

\begin{algorithm}
\begin{algorithmic}
  \STATE \textbf{input:}
  \STATE \hspace{3 mm} node $n$ in segmentation tree
  \STATE \hspace{3 mm} master mask $m_m$
  \STATE \hspace{3 mm} similarity $sim_{root}$ on root node
  \STATE \hspace{3 mm} factor $f$
  \STATE \textbf{output:}
  \STATE \hspace{3 mm} $m_m$
  \STATE
  \STATE $m_n$ := mask of $n$
  \STATE $pred\_model$ := predict attributes on image($m_n \cap m_m$)
  \STATE $sim=HI(obj\_model,pred\_model)$
  %\STATE $H_{pred}$ := predict attributes on image($m_n \cap m_m$)
  %\STATE $\displaystyle P_{pred} = \frac{H_{pattern} \cap H_{pred}}{H_{pattern} \cup H_{pred}}$
  \IF{$sim < f sim_{root}$}
    \FOR{$i=1$ to $n.childNum$}
      \STATE s\{i\}=SUBTRACTIVE\_PRUNE(n.children\{i\},$m_m$,
      \STATE ...$sim_{root}$, $f$)
    \ENDFOR
    %\STATE SUBTRACTIVE\_PRUNE(n.children)
  \ELSE
    \STATE $m_m = m_m \cap m_n$
  \ENDIF
\end{algorithmic}
\caption{SUBTRACTIVE\_PRUNE}
\label{alg:sub_prune}
\end{algorithm}

The algorithm begins at the root of the tree and predicts attributes.
This attribute distribution and its associated similarity $sim_{root}$
to the object's canonical distribution is taken to be
the standard for the image.  That is, any final segmentation should not have
significantly less similarity $sim$ than $sim_{root}$.  The algorithm traverses
the subtractive tree depth-first, predicting on each node.
Any node that doesn't have a significant negative effect on the similarity as
compared to the standard is removed.  If the
similarity was significantly affected, then we know that the region
must remain.  But it doesn't mean that sub-regions might not be able to be pruned
out, so the procedure is called recursively, checking for smaller regions
that might be able to be pruned.

If the probability was not significantly affected, then the mask is applied
and no recursive call is made.  This is important, because it means
that we can remove large chunks of the image wholesale and bypass predicting
on the node's posterity.

@@

The runtime performance of the algorithm is dominated by the prediction step.
Thus the advantage of pruning large chunks early is highly beneficial to
the runtime of the algorithm.

@@

>>>>

We use the segmentation approach of \cite{arbelaez09}.  This
produces a hierarchical segmentation by first detecting contours
using the \emph{gPb} detector \cite{maire08}.  Once the contours are
detected, regions are discovered using a watershed algorithm after
which an Ultrametric Contour Map (UCM) is constructed, which defines
the region hierarchy.  The UCM is stored as a grayscale image with
contours around the segmented regions.  These contours have intensities
representative of the region's location in the hierarchy.  So to find
the first branch in the tree, threshold the UCM image with a value
$t$ where $t = max{I(x,y)}$ where $I(x,y)$ is the intensity at pixel
location $(x,y)$.  The thresholded image can then be flood-filled into
regions.  These regions correspond to the branches on either side of
the root node in the segmentation tree.  The next level of the tree is
found by thresholding the UCM at the next highest intensity and so on.
Figure \ref{fig:ucm} shows an example UCM image and figure \ref{fig:tree}
shows the hierarchy generated from that UCM.

After the construction of segmentation tree, we detect all
the attributes in each node, including the intermediate and
leaf nodes. We use the output of SVM as the confidence (not binary),
so each node can be represented by the confidence distribution of
all attributes.  These distributions are stored and later used in
the additive pruning algorithm.  Unfortunately the subtractive algorithm
requires runtime predictions, and thus cannot use this preprocessing
step.

@@

Our objective is to detect objects and segment them out. In our
segmentation tree, the problem changes to how to select
those nodes belong to the desired object. We estimate the
probability that a node is belong to object based on the attribute
confidence distribution. For each object, we manually give the
desired attribute distribution by common knowledge. We use the
histogram intersection between the desired distribution and the
estimated confidence distribution in each node as the similarity
measure. The histogram intersection is calculated as follows:

@@

\begin{figure}
\centering
\includegraphics[width=0.30\textwidth]{figures/2008_000052_ucm.bmp.eps}
\caption{A UCM image from which a segmentation tree is derived.}
\label{fig:ucm}
\end{figure}

\begin{figure}
\centering
%\subfigure[]{
%\label{fig:treea}
\includegraphics[width=0.45\textwidth]{figures/tree_additive.eps}
%}
%\subfigure[]{
%\label{fig:treeb}
%\includegraphics[width=0.45\textwidth]{figures/tree_subtractive.eps}
%}
\caption{The segmentation hierarchy used by the additive
and subtractive algorithms.  This tree was generated from the UCM
in figure \ref{fig:ucm}.}
\label{fig:tree}
\end{figure}

@@

This pruning algorithm is based on three observations. (1) At least
one node in the segmentation tree contrains the whole object. (This
is always true since the root node contains the whole image). (2) If
both parent node and child node have high confidence of containing
the object, the segmentation of the child node is more accurate than
that of the parent node. (3) Even if some node has high confidence
of containing the object, its segmentation can still be improved by
removing irrelevant children node. The details of the additive
algorithm is shown in algorithm \ref{alg:add_prune}.

@@

\begin{algorithm}
\begin{algorithmic}
  \STATE \textbf{input:}
  \STATE \hspace{3 mm} node $n$ in segmentation tree
  \STATE \hspace{3 mm} master mask $m_m$
  \STATE \hspace{3 mm} similarity $sim_{root}$ on root node
  \STATE \hspace{3 mm} factor $f$
  \STATE \textbf{output:}
  \STATE \hspace{3 mm} mask $m_m$
  \STATE
  \STATE $m_n$ := mask of $n$
  \STATE $r = m_{m} \cap \verb+~+ m_{n}$
  \STATE $pred\_model$ := predict attributes on image($r$)
  \STATE $sim=HI(obj\_model,pred\_model)$
  %\STATE $H_{pred}$ := predict attributes on image($m_n \cap m_m$)
  %\STATE $\displaystyle P_{pred} = \frac{H_{pattern} \cap H_{pred}}{H_{pattern} \cup H_{pred}}$
  \IF{$sim < f sim_{root}$}
    \FOR{$i=1$ to $n.childNum$}
      \STATE s\{i\}=SUBTRACTIVE\_PRUNE(n.children\{i\},$m_m$,
      \STATE ...$sim_{root}$, $f$)
    \ENDFOR
    %\STATE SUBTRACTIVE\_PRUNE(n.children)
  \ELSE
    \STATE $m_m = m_m \cap \verb+~+ m_n$
  \ENDIF
\end{algorithmic}
\caption{SUBTRACTIVE\_PRUNE}
\label{alg:sub_prune}
\end{algorithm}

The algorithm begins at the root of the tree and predicts attributes.
This attribute distribution and its associated similarity $sim_{root}$
to the object's canonical distribution is taken to be
the standard for the image.  That is, any final segmentation should not have
significantly less similarity $sim$ than $sim_{root}$.  The algorithm traverses
the subtractive tree depth-first.  At each node, attributes are predicted for 
a region $r$.  
$r$ is obtained by taking the intersection of the master mask and the inverted
node mask ($r = m_{m} \cap \verb+~+ m_{n}$).  The intuition is this: the master mask is our current best guess at
the region that contains an object.  By taking the intersection, we effectively
subtract the node's region out.
We can then predict on the new region.  If 
there is no significant negative effect on the similarity as
compared to the standard, then we assume that we can safely prune off the node's
region without removing a part of the object..  If the
similarity was significantly affected, then we assume that the region
contains some part of the object and we allow it to remain.  But that 
doesn't mean that sub-regions might not be able to be pruned
out, so the procedure is called recursively, checking for smaller regions
that might be able to be pruned.

As mentioned, if the similarity score was not significantly affected by
predicting on the master mask with the region removed, then the region is
permanently removed from the master mask
and no recursive call is made.  This is important, because it means
that we can remove large chunks of the image wholesale and bypass predicting
on the node's posterity.
The runtime performance of the algorithm is dominated by the prediction step.
Thus the advantage of pruning large chunks early is highly beneficial to
the runtime of the algorithm.

@@

The subtractive algorithm knows nothing about the number of objects present
in the image.  It attempts to segment any object out into the final region.
This is evident in figure \ref{fig:sub1}.  The mobile homes in the background
have car attributes and thus are included in the final segmentation, even
though they take a far less prominent role in the unsegmented image.  Another
example is figure \ref{fig:truck}, where the two cars visually behind the truck
are segmented alongside the truck.

@@

\begin{figure}
\centering
\subfigure[]{
\label{fig:sub1a}
\includegraphics[width=0.22\textwidth]{figures/results/2008_000052.jpg.eps}
}
\subfigure[]{
\label{fig:sub1b}
\includegraphics[width=0.22\textwidth]{figures/2008_000052_1.00.jpg.eps}
} \caption{Subtractive segmentation on an image with multiple objects.
An observer may not even notice the trailer homes in the background in
the original image \subref{fig:sub1a},
but the segmentation includes a portion of them in the final segmentation \subref{fig:sub1b}.
Subtractive segmentation attempts to segment all objects present
in the image.}
\label{fig:sub1}
\end{figure}
